Rotating Stars from \Kepler Observed with Gaia DR2

### 2018-06-29 paper/ms.tex
paper full draft finished

%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

@@

Unfortunately the \Kepler survey only covers a single pointing, and so our volume analysis of the stellar ages is limited here. However, with more than 15 distinct lines of sight available from K2, the extended \Kepler mission, we will soon be able to more than double our sample of rotating field stars and probe a much wider range of stellar ages. This may enable us to determine the spatial scales that star formation histories are coherant over in the nearby Milky Way, and make comparisons to the $\sim$100 pc resolved star formation history maps available for nearby galaxies such as Andromeda \citep[][]{lewis2015,williams2017}. For example, gyrochronology ages for stars in disk within 1 kpc of the Sun could be used to test if star formation propagates due to spiral arm density waves \citep[e.g. see][]{choi2015}.

While we have removed as contaminants the prominent parallel main sequence of nearly equal-mass binary stars in this analysis, as well as sub-giant stars, both of these samples are clearly deserving of further analysis. %% HERE

finally, this work has found strong motivation for a new generation of gyrochronology isochrones to adequately model the {\em Gaia}--\Kepler/K2 (and soon TESS) rotation data for field stars. As we show in Figure \ref{fig:cmd_zoom}, these models should include both the angular momentum evolution (period), as well as the temperature and luminosity (color, magnitude) as a function of time. With such improvements, we may be able to

>>>>

\clearpage
%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

@@

Unfortunately the \Kepler survey only covers a single pointing, and so our volume analysis of the stellar ages is limited here. However, with more than 15 distinct lines of sight available from K2, the extended \Kepler mission, we will soon be able to more than double our sample of rotating field stars and probe a much wider range of stellar ages. This may enable us to determine the spatial scales that star formation histories are coherant over in the nearby Milky Way, and make comparisons to the $\sim$100 pc resolved star formation history maps available for nearby galaxies such as Andromeda \citep[][]{lewis2015,williams2017}. For example, gyrochronology ages for disk stars within 1 kpc of the Sun could be used to test if star formation propagates due to spiral arm density waves \citep[e.g. see][]{choi2015}.

While we have removed as contaminants the prominent parallel main sequence of nearly equal-mass binary stars in this analysis, as well as sub-giant stars, both of these samples are clearly deserving of further analysis. For example, comparing the rotation period distribution for binary versus single main sequence stars may provide constraints on their dynamical histories.

Finally, in the {\em Gaia} era, stellar evolution modelers have the daunting task to explain an increasing number of precisely determined features from survey observations, such as the fascinating new main sequence gap revealed by \citet{jao2018}, and the main sequence width seen here in Figure \ref{fig:cmd}.
This work has presented strong motivation for a new generation of stellar evolution and rotation models to be produced to accurately model the {\em Gaia} main sequence. With \Kepler, K2, and soon TESS, rotation measurements will be available for hundreds of thousands of nearby stars that {\em Gaia} has provided precise distances for. We hope new isochrone models will reproduce both the observed colors and luminosities of stars due to their ages and abundances, as well as the evolution of their surface rotations.
